A Simple Cache-Sensitive Skip List (SCSSL)

### 2025-05-27 paper.tex
affiliation info
test
add affiliation
state should be country
suggestions to paper \[not complete\]

\newcommand{\keywords}[1]{\par\noindent\textbf{Keywords:} #1}
% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{tikz}
\usepackage{pgfplots}
\usetikzlibrary{arrows.meta}
% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\SetKw{KwDownTo}{down to}

\title{simple cache-sentive skip list}
\author{Yuchang Ke \\
 \texttt{aceking.ke@gmail.com}
\and Bill Lv \\
\texttt{ideaalloc@gmail.com}
\and Weijie Zhang\\
\texttt{zhangweijie@gmail.com}}

@@

\begin{abstract}
      We present a cache-sensitive skiplist that is very simple and easy to implement. It only requires a few small modifications to the original skiplist to adapt it to modern CPU architectures, resulting in a performance improvement of 1.2~1.6 times, in some system it can reach 2X or 3X.
      Unlike other optimizations, it does not require specialized instructions such as SIMD or multi-threaded programming. Instead, it takes advantage of modern compiler optimizations, making it more universal. This design can be used not only in in-memory databases or key-value stores, but also in embedded systems.

\keywords{skiplist; cache-sensitive; index structure}
\end{abstract}

\section{Introduction}

In the field of computer science, ordered data structures are a fundamental tool for organizing and managing data. Efficient execution of search, delete, and insert operations is crucial for a wide range of applications, as it directly impacts overall system performance.
To achieve high performance in ordered data structures, many methods have been developed, such as AVL trees and Red-Black trees. However, these structures are relatively complex due to the need for rebalancing. This changed when William Pugh introduced the skip list in 1990\cite{ref1}. It introduced a randomized data structure that simplified implementation compared to traditional balanced trees.

Thanks to its simplicity, the skip list has been adopted by various systems, including databases such as LevelDB\cite{ref2}, Redis\cite{ref7}, and RocksDB\cite{ref3}, Hbase\cite{ref4} in-memory caches, and the inverted index in Lucene.

Skip list can be divided into multiple levels, one data level and multiple index levels. Data level is the lowest level(level 0), which stores all the data nodes. And the index levels provide short paths to access the data node at the data level, and all the levels are linked lists. Because of this, these data nodes and index nodes are non-contiguous in memory address. This causes the performance to be degraded by cache misses.
Although ESL, CSSL has developed cache sensitive skip lists to fit modern CPU architecture, these algorithms are still complex, and CSSL uses SIMD instructions which are special in Intel. We strongly believe that the value of a skiplist lies in its simplicity.  Excessively complex cache-sensitive designs reduce the appeal of skip list variations.

With the emergence of new technologies and computing architectures, particularly multi-core CPUs and vectorization, novel variants of skip lists have been developed. Such as Cache-Sensitive Skiplist (CSSL)\cite{ref5}, which merges nodes at the index level into a single array, which are named fast lanes, and all nodes in the fast lane are sorted. This design is more cache-friendly. It also uses SIMD instructions, ESL( A High-Performance Skiplist with Express Lane\cite{ref6})  also merges the nodes in index levels into a single array, and they use version-based lock protocol and lock-free operation log to implement concurrency.

\section{Design overview}

In this paper, we present a simple and easy-to-implement cache-sensitive skip list. At the data level, it introduces a vector, and the nodes at this level now store arrays instead of single values. Leveraging CPU cache benefits for small arrays to enhance skip list performance. For example, consider a skip list with a vector at the data level; let's suppose the vector's maximum size is 2, as shown in Figure 1.

There are 3 key points: 1) At the data level, we use a sorted array with a fixed maximum length. Since the keys are integers, we choose a maximum length of 128 or less. 2) For all index levels, we use only the first element of the array as a pivot for comparison. 3) The last element of the array is considered solely during insertion operations, and only under particular conditions.

\begin{figure}[h]
\centering
\includegraphics[width=0.5\linewidth]{skiplist.png}
\caption{A simple cache-sensitive skip list}
\label{fig:fig1}
\end{figure}
Our key contributions include: 1. Introducing a vector, which maintains the algorithm's complexity while improving CPU cache hit rate due to its contiguous storage. 2. Reducing the skiplist's level count and pointer overhead by incorporating a vector. 3. Presenting vector-based insertion, search, and deletion algorithms for data nodes, offering a novel perspective for database index design.

\section{Algorithm}
This section gives the algorithm for search, insertion, and deletion. The insertion operation inserts new specified data into the data structure, and the search operation finds data and returns it. The deletion operation deletes data.

In SCSSL, at the data level, data is stored in a single array with a fixed maximum size. As the same as the standard skip list, the level of the node is chosen randomly when the node is created. And a level i node has the same number of forward pointers (also i), the maximum of levels is a constant MaxLevel. It also has a header that leads the search of the data structure, and the last nodes of all levels’ forward pointers point to NIL.

\begin{figure}[ht]
      \centering
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,302); %set diagram left start at 0, and has height of 302

@@

\draw [shift={(153.04,119.22)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(248.29,117.98)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(334.01,116.13)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(454.37,117.36)}, rotate = 180.58] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(540.96,116.75)}, rotate = 181.36] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(153.91,93.91)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(153.91,69.23)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(538.36,67.38)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(332.28,92.06)}, rotate = 180.3] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(540.96,91.45)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(425.8,51.33)}, rotate = 317.09] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(157.37,235.44)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(252.62,234.21)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(338.34,232.36)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(416.27,233.7)}, rotate = 182.18] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(583.39,232.97)}, rotate = 181.36] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(158.24,210.14)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(158.24,185.46)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(580.79,183.61)}, rotate = 179.94] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(336.61,208.29)}, rotate = 180.3] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(583.39,207.67)}, rotate = 180.02] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

@@

\draw [shift={(495.07,233.7)}, rotate = 182.18] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

% Text Node
\draw (170.55,108.56) node [anchor=north west][inner sep=0.75pt]   [align=left] {3,4};
% Text Node
\draw (264.06,109.17) node [anchor=north west][inner sep=0.75pt]   [align=left] {5,6};
% Text Node
\draw (349.79,107.94) node [anchor=north west][inner sep=0.75pt]   [align=left] {7,9};
% Text Node
\draw (470.74,107.94) node [anchor=north west][inner sep=0.75pt]   [align=left] {12,17};
% Text Node
\draw (174.82,85.72) node [anchor=north west][inner sep=0.75pt]   [align=left] {3};
% Text Node
\draw (177.41,60.42) node [anchor=north west][inner sep=0.75pt]   [align=left] {3};
% Text Node
\draw (352.32,82.64) node [anchor=north west][inner sep=0.75pt]   [align=left] {7};
% Text Node
\draw (76.23,40.06) node [anchor=north west][inner sep=0.75pt]   [align=left] {head};
% Text Node
\draw (553.41,39.44) node [anchor=north west][inner sep=0.75pt]   [align=left] {tail};
% Text Node
\draw (564.06,59.8) node [anchor=north west][inner sep=0.75pt]   [align=left] {nil};
% Text Node
\draw (564.06,85.72) node [anchor=north west][inner sep=0.75pt]   [align=left] {nil};
% Text Node
\draw (562.33,106.7) node [anchor=north west][inner sep=0.75pt]   [align=left] {nil};
% Text Node
\draw (45.5,20.51) node [anchor=north west][inner sep=0.75pt]   [align=left] {search path};
% Text Node
\draw (435.12,22.78) node [anchor=north west][inner sep=0.75pt]   [align=left] {update[i] →forward[i] };
% Text Node
\draw (265.32,141.26) node [anchor=north west][inner sep=0.75pt]   [align=left] {original list, 8 to be inserted};
% Text Node
\draw (174.88,224.78) node [anchor=north west][inner sep=0.75pt]   [align=left] {3,4};
% Text Node
\draw (268.39,225.4) node [anchor=north west][inner sep=0.75pt]   [align=left] {5,6};
% Text Node
\draw (354.12,224.17) node [anchor=north west][inner sep=0.75pt]   [align=left] {7,8};
% Text Node
\draw (505.37,224.17) node [anchor=north west][inner sep=0.75pt]   [align=left] {12,17};
% Text Node
\draw (179.15,201.95) node [anchor=north west][inner sep=0.75pt]   [align=left] {3};
% Text Node
\draw (181.74,176.65) node [anchor=north west][inner sep=0.75pt]   [align=left] {3};
% Text Node
\draw (356.65,198.86) node [anchor=north west][inner sep=0.75pt]   [align=left] {7};
% Text Node
\draw (80.56,156.28) node [anchor=north west][inner sep=0.75pt]   [align=left] {head};
% Text Node
\draw (595.83,155.67) node [anchor=north west][inner sep=0.75pt]   [align=left] {tail};
% Text Node
\draw (606.49,176.03) node [anchor=north west][inner sep=0.75pt]   [align=left] {nil};
% Text Node
\draw (606.49,201.95) node [anchor=north west][inner sep=0.75pt]   [align=left] {nil};
% Text Node
\draw (604.76,222.93) node [anchor=north west][inner sep=0.75pt]   [align=left] {nil};
% Text Node
\draw (431.99,223.16) node [anchor=north west][inner sep=0.75pt]   [align=left] {9};
% Text Node
\draw (267.41,246.44) node [anchor=north west][inner sep=0.75pt]   [align=left] {list after insertation};

\end{tikzpicture}
\caption{insert 8 into simple cache-sentive skip list(max array size =2)}
\label{fig:fig2}
\end{figure}

\subsection{Initialization}
The initialization operator is elementary, just create a header with MaxLevel forwards array, and set them to Nil. To simplify the algorithm, we set the header's array with a single value, -inf.

>>>>

% Language setting
\usepackage[english]{babel}

% Set page size and margins
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{booktabs} % For professional tables
\usepackage{pgfplots} % For plots
\pgfplotsset{compat=1.18} % Recommended for current pgfplots versions

\newcommand{\keywords}[1]{\par\noindent\textbf{Keywords:} #1}
\SetKwDownTo{KwDownTo}{down to} % Corrected definition from original to ensure it works
\SetKwFor{For}{for}{do}{endfor}
\SetKwIF{If}{ElseIf}{Else}{if}{then}{else if}{else}{endif}
% algorithm2e setup for NIL
\SetKw{Nil}{NIL}

\title{A Simple Cache-Sensitive Skip List (SCSSL)}
\author{Yuchang Ke \\
\texttt{aceking.ke@gmail.com}
\and Bill Lv \\
\texttt{ideaalloc@gmail.com}
\and Weijie Zhang\\
\texttt{zhangweijie@gmail.com}}
\date{\today}

@@

\begin{abstract}
We present a cache-sensitive skip list (SCSSL) that is remarkably simple and easy to implement. It requires only minor modifications to the original skip list data structure to adapt it effectively to modern CPU architectures. This adaptation results in a performance improvement typically ranging from 1.2x to 1.6x, and on some systems, can reach 2x to 3x for insertion-heavy workloads.
Unlike other optimization techniques, SCSSL does not rely on specialized hardware instructions like SIMD or intricate multi-threaded programming paradigms. Instead, its design inherently benefits from improved data locality due to the use of arrays at the data level, a feature that modern compilers can further leverage. This makes the proposed design broadly applicable, suitable not only for in-memory databases and key-value stores but also for resource-constrained embedded systems.
\keywords{skip list; cache-sensitive; index structure; data structures; performance optimization}
\end{abstract}

\section{Introduction}
\label{sec:introduction}

In computer science, ordered data structures are fundamental for organizing and managing data efficiently. The performance of search, insertion, and deletion operations within these structures is crucial for a wide array of applications, as it directly impacts overall system performance.
Numerous methods have been developed to achieve high performance in ordered data structures, such as AVL trees and Red-Black trees. However, these balanced tree structures are relatively complex to implement due to the necessity of rebalancing operations. This landscape changed when William Pugh introduced the skip list in 1990 \cite{Pugh1990}. The skip list, a probabilistic data structure, offered a simpler implementation alternative to traditional balanced trees while providing comparable performance characteristics.

Thanks to its simplicity and efficiency, the skip list has been widely adopted in various systems. Notable examples include databases like LevelDB \cite{LevelDB}, Redis \cite{Redis}, RocksDB \cite{RocksDB}, and HBase \cite{HBase}, as well as in-memory caches and the inverted index component of search engines like Lucene.

A standard skip list consists of multiple levels: a data level (level 0) that stores all data nodes in a sorted linked list, and multiple index levels above it. These index levels contain a subset of the nodes from lower levels and act as "express lanes" to expedite traversal to the desired data nodes. However, nodes in a skip list are typically allocated dynamically, leading to non-contiguous memory addresses. This fragmentation can degrade performance on modern CPUs due to frequent cache misses.

Although cache-sensitive skip list variants like the Cache-Sensitive Skip List (CSSL) \cite{CSSL20XX} and ESL (Express Lane Skip List) \cite{ESL20XX} have been developed to better align with modern CPU architectures, they often introduce significant complexity. For instance, CSSL utilizes SIMD instructions, which may not be universally available or optimal across all platforms. ESL proposes merging index level nodes into arrays and employs sophisticated version-based locking protocols and lock-free operation logs for concurrency. We believe that the primary appeal of the skip list lies in its inherent simplicity. Overly complex cache-sensitive designs can diminish this advantage.

Our work is motivated by the observation that simple architectural changes can yield substantial performance gains by improving cache locality. We propose a modification that retains the core simplicity of the skip list while significantly enhancing its cache-friendliness.

\section{Design Overview}
\label{sec:design_overview}

In this paper, we present a Simple Cache-Sensitive Skip List (SCSSL). The core idea is to modify the nodes at the data level (level 0) to store a small, sorted array of elements (e.g., key-value pairs or just keys) instead of a single element. This approach leverages CPU cache locality for these small arrays, thereby enhancing performance. For example, Figure \ref{fig:conceptual_scssl} illustrates a conceptual SCSSL where data nodes hold arrays of maximum size 2.

The design of SCSSL hinges on three key principles:
\begin{enumerate}
    \item \textbf{Array-based Data Nodes:} At the data level (level 0), each node contains a sorted array with a fixed maximum length. For integer keys, a maximum length that allows the node to fit well within CPU cache lines (e.g., up to 128 elements, depending on key/value sizes) is chosen.
    \item \textbf{Index Level Pivot:} For navigation through the index levels (levels 1 and above), only the first element (i.e., the smallest key) of the array in a data-level node (or an index node pointing to such an array) is used as the pivot for comparison.
    \item \textbf{Insertion Handling:} The last element (i.e., the largest key) of an array in a data node is primarily considered during insertion operations, particularly when an array is full and a new key needs to be accommodated.
\end{enumerate}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.7\linewidth]{skiplist.png} % Assuming skiplist.png exists
\caption{A conceptual view of the Simple Cache-Sensitive Skip List (SCSSL) where data nodes (level 0) store arrays of elements (e.g., max array size = 2 as depicted). Index nodes point to the start of these arrays.}
\label{fig:conceptual_scssl}
\end{figure}

Our key contributions include:
\begin{enumerate}
    \item Introducing arrays within data-level nodes, which maintains the asymptotic complexity of skip list operations while significantly improving CPU cache hit rates due to the contiguous storage of multiple elements.
    \item Reducing the effective height of the skip list and pointer overhead, as each data-level node now stores multiple elements, leading to fewer nodes overall for a given dataset size.
    \item Presenting modified insertion, search, and deletion algorithms tailored for these array-based data nodes, offering a practical and simple approach for enhancing skip list performance, particularly in database indexing and similar applications.
\end{enumerate}

\section{Algorithms}
\label{sec:algorithms}
This section details the algorithms for search, insertion, and deletion in SCSSL. These operations are adapted from standard skip list algorithms to accommodate the array-based nodes at the data level.

In SCSSL, each node at the data level (level 0) stores its elements in a sorted array of a fixed maximum size. When a new node is created, its level is chosen randomly, similar to a standard skip list. A node promoted to level $k$ has $k+1$ forward pointers (indexed 0 to $k$). The maximum number of levels in the skip list is a predefined constant, `MAX_POSSIBLE_LEVEL`. The skip list employs a header node to initiate searches; its forward pointers are initially \Nil. For simplicity in comparisons, the header node can be conceptually considered to hold a key value of negative infinity or have a special marker.

Figure \ref{fig:insertion_example} illustrates an example of an insertion operation in SCSSL.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]

@@

\draw [shift={(153.04,119.22)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(248.29,117.98)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(334.01,116.13)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(454.37,117.36)}, rotate = 180.58] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(540.96,116.75)}, rotate = 181.36] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(153.91,93.91)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(153.91,69.23)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(538.36,67.38)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(332.28,92.06)}, rotate = 180.3] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(540.96,91.45)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(425.8,51.33)}, rotate = 317.09] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(157.37,235.44)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(252.62,234.21)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(338.34,232.36)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(416.27,233.7)}, rotate = 182.18] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(583.39,232.97)}, rotate = 181.36] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(158.24,210.14)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(158.24,185.46)}, rotate = 180] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(580.79,183.61)}, rotate = 179.94] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(336.61,208.29)}, rotate = 180.3] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(583.39,207.67)}, rotate = 180.02] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

@@

\draw [shift={(495.07,233.7)}, rotate = 182.18] [color={rgb, 255:red, 0; green, 0; blue, 0 }][line width=0.75] (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29) ;

% Text Node
\draw (170.55,108.56) node [anchor=north west][inner sep=0.75pt] [align=left] {3,4};
% Text Node
\draw (264.06,109.17) node [anchor=north west][inner sep=0.75pt] [align=left] {5,6};
% Text Node
\draw (349.79,107.94) node [anchor=north west][inner sep=0.75pt] [align=left] {7,9};
% Text Node
\draw (470.74,107.94) node [anchor=north west][inner sep=0.75pt] [align=left] {12,17};
% Text Node
\draw (45.5,20.51) node [anchor=north west][inner sep=0.75pt] [align=left] {Search path for 8};
% Text Node
\draw (435.12,22.78) node [anchor=north west][inner sep=0.75pt] [align=left] {update[i]->forward[i]};
% Text Node
\draw (265.32,141.26) node [anchor=north west][inner sep=0.75pt] [align=left] {Original list, key 8 to be inserted};
% Text Node
\draw (174.88,224.78) node [anchor=north west][inner sep=0.75pt] [align=left] {3,4};
% Text Node
\draw (268.39,225.4) node [anchor=north west][inner sep=0.75pt] [align=left] {5,6};
% Text Node
\draw (354.12,224.17) node [anchor=north west][inner sep=0.75pt] [align=left] {7,8}; % Node modified
% Text Node
\draw (505.37,224.17) node [anchor=north west][inner sep=0.75pt] [align=left] {12,17};
\draw (431.99,223.16) node [anchor=north west][inner sep=0.75pt] [align=left] {9}; % New node contains 9
% Text Node -- Corrected typo from original paper
\draw (267.41,246.44) node [anchor=north west][inner sep=0.75pt] [align=left] {list after insertion};
\end{tikzpicture}
\caption{Insertion of key 8 into an SCSSL (max array size = 2). Key 8 displaces 9 from node (7,9). Key 9 is then inserted into a new node.}
\label{fig:insertion_example}
\end{figure}

\subsection{Initialization}
The initialization process is elementary. A header node is created with an array of `MAX_POSSIBLE_LEVEL + 1` forward pointers, all initialized to \Nil. To simplify comparisons, the header node's key array (if it were to have one in the same structure) can be considered to contain a single sentinel value of negative infinity, or it can be handled as a special case. In our algorithms, `header.FirstKey()` would effectively be $-\infty$.

\subsection{Search}
The search algorithm (Algorithm \ref{alg:search}) navigates the skip list from the highest level down to level 0. At each level `i`, it traverses forward as long as the first key of the array in the next node (`current.forward[i].FirstKey()`) is less than or equal to the search key `k`. Once this traversal is complete, the `current` node is the one whose own array might contain `k` (specifically, `current.FirstKey() <= k`, and either `current.forward[0]` is \Nil or `current.forward[0].FirstKey() > k`). The final step involves searching for `k` within `current`'s array.

\begin{algorithm}[H]
    \caption{Search for Key $k$ in SCSSL}
    \label{alg:search}
    \KwIn{Search key $k$, current global maximum level `currentGlobalMaxLevel`, header node `header`}
    \KwOut{Node containing key $k$ and its index in array, or (\Nil, -1) if not found}
    \SetKwFunction{FSearch}{Search}
    \SetKwProg{Fn}{Function}{:}{}
    \Fn{\FSearch{$k$, currentGlobalMaxLevel, header}}{
        $current \gets header$\;
        \For{$i \gets \text{currentGlobalMaxLevel}$ \KwDownTo $0$}{
            \While{$current.forward[i] \neq \Nil \land current.forward[i].FirstKey() \leq k$}{
                $current \gets current.forward[i]$\;
            }
        }
        \Return $current.\text{search\_in\_array}(k)$\; \tcp{Returns (node, index) or (\Nil, -1)}
    }
\end{algorithm}

\subsection{Insertion}
The insertion algorithm (Algorithm \ref{alg:insert}) first searches for the appropriate position for the new key `k`, maintaining an `update` array to store pointers to the predecessor nodes at each level. The core logic then addresses several cases based on whether the target data node (`target_node`, which is `update[0]`) can accommodate `k`:
